---
title: "College Majors & Salary: Risk vs. Return?"
author: "Talha Siddiqui"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(scales)
library(ggrepel)
library(broom)
library(plotly)
theme_set(theme_light())
```

```{r read CSV}
recent_grad <- read.csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018-10-16/recent-grads.csv")

majors_processed <- recent_grad %>% 
  arrange(desc(Median)) %>% 
  mutate(
    Major = str_to_title(Major),
    Major = fct_reorder(Major,Median)
  )
```


```{r by_major_category}
by_major_category <- majors_processed %>% 
  filter(!is.na(Total)) %>% 
  group_by(Major_category) %>% 
  summarise(
    Men = sum(Men),
    Women = sum(Women),
    Total = sum(Total),
    MedianSalary = sum(Median * Sample_size)/sum(Sample_size)
  ) %>% 
  mutate(ShareWomen = Women / Total) %>% 
  arrange(desc(ShareWomen))
```

I am very excited to share a cool learning opportunity that I came across last week. [David Robinson conducted a great screencast walking through a live data analysis in R.](http://varianceexplained.org/r/tidy-tuesday-college-major/) It was a terrific experience watching an expert share their approach of quickly analyzing a dataset. I followed along David's exploratory data analysis and plotting (admittedly by pausing, rewinding and catching up several times).

Following his model, I wanted the data to answer a question of my own, *does pursing a higher salary major risk in higher chances of unemployment?* Let's take a look:

## Risk vs. Return? Or better, Both!

```{r}
ggplotly(
  majors_processed %>% 
    mutate(Major_category = fct_lump(Major_category, 5)) %>% 
    filter(Sample_size >= 100) %>% 
    ggplot(aes(x = Median, y = Unemployment_rate, label = Major)) +
    geom_point(aes(colour = Major_category, size = Sample_size)) +
    expand_limits(y = 0) + 
    scale_y_continuous(labels = percent_format()) +
    scale_x_continuous(labels = dollar_format()) +
    coord_flip() +
    labs(
      title = "Risk vs. Return? Or Both?",
      subtitle = "Looking at the relationship of Median income and unemployment",
      x = "Median Income of Recent Graduates",
      y = "Unemployment in Recent Graduates",
      legend = "Major Category"
    )
)
```

**It appears that you can have your cake and eat it too.** Higher earning majors do get to enjoy lower unemployment, but if unemployment is your only concern then Engineering isn't your only option. There are plenty of majors that will guarantee you job security but are not as financially rewarding.

The following analysis is my effort at learning and following along David's analysis. If you're interested, check out [the github repository containing the R Markdown file](https://github.com/talsiddiqui/recentgrads).

### Most common majors

#### What are the most common major categories (e.g. Business, Engineering, Social Sciences etc.)?

```{r}
by_major_category %>% 
  mutate(Major_category = fct_reorder(Major_category, Total)) %>% 
  gather(Gender, Number, Men, Women) %>% 
  ggplot(aes(Major_category, Number, fill = Gender)) + 
  geom_col() +
  coord_flip() +
  labs(
    title = "Most common Major Categories",
    x = "",
    y = "No. of Total Graduates"
  ) + 
  scale_y_continuous(labels = comma_format())
```


### What categories of majors make more money than others?

```{r}
recent_grad %>% 
  ggplot(aes(fct_reorder(Major_category,Median), Median, fill = Major_category)) +
  geom_boxplot() +
  scale_y_continuous(labels = dollar_format()) +
  coord_flip() +
  expand_limits(y = 0) +
  labs(
    title = "Highest Earning Major Categories",
    x = "",
    y = "Median Salaries"
  ) + 
  theme(legend.position = "none")
```


### What are the highest earning majors?
Let's only consider at least 100 surveyed graduates per major

```{r}
majors_processed %>% 
  filter(Sample_size >= 100) %>% 
  head(20) %>% 
  ggplot(aes(Major, Median, colour = Major_category)) +
  geom_point() +
  geom_errorbar(aes(ymin = P25th, ymax = P75th)) +
  expand_limits(y = 0) +
  coord_flip() +
  scale_y_continuous(labels = dollar_format()) +
  labs(
    x = "",
    title = "What are the highest earning majors?",
    subtitle = "Top 20 majors with at least 100 graduates surveyed.\nBars represent 25th to 75th percentile.",
    legend = "Major Category",
    y = "Median Salary of Recent Graduates"
  )
```


### How does gender breakdown relate to earnings? 

```{r warning = FALSE}
ggplotly(
  majors_processed %>% 
    mutate(Major_category = fct_lump(Major_category, 5)) %>% 
    ggplot(aes(ShareWomen, Median, label = Major)) +
    geom_point(aes(colour = Major_category, size = Sample_size)) +
    geom_smooth(method = "lm") +
    expand_limits(y = 0) + 
    scale_x_continuous(labels = percent_format()) +
    scale_y_continuous(labels = dollar_format()) + 
    labs(
      x = "Percentage of Women",
      title = "Gender Breakdown & Earnings",
      subtitle = "Linear model representing the relationship between\nMedian Salaries and Percentage of Women",
      legend = "Major Category",
      y = "Median Salary of Recent Graduates"
    )
)
```



## Appendix

### How good is our data? Do we have enough surveys for each major?

```{r}
majors_processed %>% 
  ggplot(aes(Sample_size, Median)) +
  geom_point() +
  geom_text(aes(label = Major), check_overlap = TRUE) +
  geom_vline(xintercept = 100, color = "red") +
  scale_x_log10() +
  labs(
    x = "No. of Surveyed Graduates",
    title = "Do we have enough surveys for each major?",
    subtitle = "There seem to be a lot of majors with too few surveyed graduates"
  )
```

### Some statistical calculations

```{r weighted linear model}
majors_processed %>% 
  select(Major, Total, ShareWomen, Sample_size, Median) %>% 
  lm(Median ~ ShareWomen, data = ., weights = Sample_size) %>% 
  summary()
```


```{r}
majors_processed %>% 
  select(Major, Major_category, Total, ShareWomen, Sample_size, Median) %>% 
  add_count(Major_category) %>% 
  filter(n >= 10) %>% 
  nest(-Major_category) %>% 
  mutate(model = map(data, ~ lm(Median ~ ShareWomen, data = ., weights = Sample_size)),
         tidied = map(model, tidy)) %>% 
  unnest(tidied) %>% 
  filter(term == "ShareWomen") %>% 
  arrange(estimate)
```


```{r}
knitr::knit_exit()
```


## Scrap Work

#### What are the most common *majors*?
(There are a total of 173, so we will check out the top 20)

```{r}
majors_processed %>% 
  mutate(Major = fct_reorder(Major, Total)) %>% 
  arrange(desc(Total)) %>% 
  head(20) %>% 
  ggplot(aes(Major, Total, fill = Major_category)) + 
  geom_col() +
  scale_y_continuous(labels = comma_format()) +
  coord_flip() +
  labs(
    x = "",
    y = "No. of Total Graduates"
  )
```

```{r}
recent_grad %>% 
  group_by(Major_category) %>% 
  summarise(Median = median(Median)) %>% 
  ggplot(aes(fct_reorder(Major_category,Median), Median)) +
  geom_col() +
  scale_y_continuous(labels = dollar_format()) +
  coord_flip()
```

### What are the highest earning majors? Without considering how many samples per major

```{r}
majors_processed %>% head(20) %>% 
  ggplot(aes(Major, Median, colour = Major_category)) +
  geom_point() +
  geom_errorbar(aes(ymin = P25th, ymax = P75th)) +
  expand_limits(y = 0) +
  coord_flip() +
  scale_y_continuous(labels = dollar_format()) +
  labs(
    x = "Median Salary of Graduates",
    title = "What are the highest earning majors?",
    subtitle = "Top 20 majors with at least 100 graduates surveyed. Bars represent 25th to 75th percentile."
  )
```



### What are the lowest earning majors?

```{r}
recent_grad %>% 
  arrange(Median) %>% 
  select(Major, Major_category, Median, P25th, P75th) %>% 
  head(20) %>% 
  mutate(
    Major = str_to_title(Major),
    Major = fct_reorder(Major,desc(Median))
  ) %>% 
  ggplot(aes(Major, Median, colour = Major_category)) +
  geom_point() +
  geom_errorbar(aes(ymin = P25th, ymax = P75th)) +
  expand_limits(y = 0) +
  coord_flip() +
  scale_y_continuous(labels = dollar_format())
```

### Gender related plots
```{r}
majors_processed %>% 
  arrange(desc(Total)) %>% 
  head(20) %>% 
  mutate(Major = fct_reorder(Major, Total)) %>% 
  gather(Gender, Number, Men, Women) %>% 
  ggplot(aes(Major, Number, fill = Gender))+
  geom_col() +
  coord_flip()
```

```{r}
by_major_category %>% 
  ggplot(aes(ShareWomen, MedianSalary)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text_repel(aes(label = Major_category), force = 0.2) + 
  expand_limits(y = 0)
```
